$Z_3$ topological order in the quantum dimer-pentamer model

### 2016-11-27 qdpm_paper.tex
some fixes

We can define two oriented open string operators which commute with $H_{Z_3}$ everywhere except at the endpoints. The electric string operator $S_e$, which acts on an open, oriented string on the lattice $\gamma_{vv'}$ with endpoints $v,v'$, and creates a $+1$ ($-1$) charge at vertex $v$ ($v')$, is defined as
\begin{equation}
    S_e \left( \gamma_{vv'} \right) \equiv \prod_{l \in \gamma_{vv'}} Q_{v_l v_l'}^\dagger
\end{equation}
where the product is over all links $l$ in $\gamma_{vv'}$, $v_l$ and $v_l'$ are the two vertices in $l$, ordered according to the orientation of $\gamma_{vv'}$ (see \figref{fig:example_elec_string}). The magnetic string operator $S_m$ is defined by an open, oriented string on the dual lattice and creates a magnetic vortex-antivortex pair on plaquettes at the endpoints of $\tilde{\gamma}_{pp'}$
\begin{equation}
    S_m \left( \tilde{\gamma}_{vv'} \right) \equiv \prod_{l \perp \tilde{\gamma}_{pp'}} P_{v_l v_l'}^\dagger
\end{equation}
where the product is over all links $l$ that cross $\tilde{\gamma}$ and $v_l$ ($v_l'$) is the vertex to the left (right) of the string as $\tilde{\gamma}$ is traversed from $p$ to $p'$. 
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
    \centering
    \includegraphics[width=0.8\linewidth]{example_elec_string.pdf}
    \caption{ ({\it left}) An example of a charge free $Z_3$ electric string configuration. The electric string operator $S_\gamma^e$ represented by the thick gray line, oriented from the red to blue vertex. ({\it right}) An example of a $Z_3$ electric string configuration with a $+1$ ($-1$) electric charge at the red (blue) vertex; this string configuration results form the action of $S_\gamma^e$ on the left panel. \note{Owen: Can you remove the a) and b) labels on the figure? Also switch the $\Gamma$ to $\gamma$, and remove the $v_0-v_3$ labels. Add a gray arrow on the string operator from red to blue. Can you add a non-trivial magnetic loop around the blue charge on the right?} }
    \label{fig:example_elec_string}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}
%    \centering
%    \includegraphics[width=1.0\columnwidth]{spatial_cors_perp.pdf}
%    \caption{Correlation function between perpendicular dimers on the same (blue) and different (red) 
%    sub-lattice for a $L=64$ lattice.}
%    \label{fig:spatial_dmr_cor}
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%

@@

To test for the possibility of translational and rotational symmetry breaking corresponding to the features in $S_d$ at the Brillouin zone edges, we define several order parameters which would capture such potential symmetry breaking. For rotational symmetry breaking we define 
\begin{equation}
\Mrot \equiv  \frac{2}{L^2} \left(N_h - N_v\right) \label{eq:Mrot},
\end{equation}
where $N_v$ ($N_h$) is the total number of vertical (horizontal) dimers; note that $\exval{M_{\rm{rot}}^2 } \neq 0$ implies there is rotational symmetry breaking. to To capture translational symmetry breaking by one lattice site in a direction perpendicular to the majority of the dimers we define $\Mperp$ 
\begin{align}
\Mperp \equiv  \frac{8}{L^2} \sum_{\Box} \Bigl(  &(-1)^{r_x} \theta\left(\Mrot \right) \Pi_\Box^v \notag\\
+ &(-1)^{r_y}\theta\left(-\Mrot \right) \Pi_\Box^h\Bigr) \label{eq:Mperp},
\end{align}
where $\theta(x)$ is the Heaviside function, and $\Pi_\Box^h$ ($\Pi_\Box^v$) is one if the plaquette $\Box$ has two horizontal (vertical) dimers, and zero otherwise. Similarly, we define $\Mpar$ to be given by \Eqref{eq:Mperp} with $r_x$ and $r_y$ swapped between terms. Additionally, we define order parameters that indicate ``columnar" and ``staggered" dimer order~\cite{Syljuasen2005}:
\begin{align}
\MstagSq \equiv  \frac{1}{L^2} \biggl[ &\biggl( \sum_{\ell_h} (-1)^{r_x+r_y} n_d \left(\vec{r}_{\ell_h}\right) \biggr)^2  \notag\\
+&\biggl( \sum_{\ell_v} (-1)^{r_x+r_y} n_d \left(\vec{r}_{\ell_v}\right) \biggr)^2\biggr]\label{eq:Mstag},
\end{align}
\begin{align}
\McolSq \equiv  \frac{1}{L^2} \biggl[  &\biggl( \sum_{\ell_h} (-1)^{r_x} n_d \left(\vec{r}_{\ell_h}\right) \biggr)^2 \notag \\
+ &\biggl( \sum_{\ell_v} (-1)^{r_y} n_d \left(\vec{r}_{\ell_v}\right) \biggr)^2  \biggr]\label{eq:Mcol},
\end{align}
where the sums over $\ell_h$ ($\ell_v$) are taken over all horizontal (vertical) links. In \figref{fig:order_params} we have plotted the finite-size scaling of these order parameters; we find in all cases that while finite size systems display symmetry breaking due to finite-size effects, these order parameters all vanish in the thermodynamic limit, as demonstrated by the extrapolation in \figref{fig:order_params}.
%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[]
    \centering
    \includegraphics[width=1.0\linewidth]{order_params_plot.pdf}
    \caption{The symmetry breaking order parameters defined in \eqref{Mperp}.}
    \label{fig:order_params}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

>>>>

We can define two oriented open string operators which commute with $H_{Z_3}$ everywhere except at the endpoints. The electric string operator $S_e$, which acts on an open, oriented string on the lattice $\gamma_{vv'}$ with endpoints $v,v'$, and creates a $+1$ ($-1$) charge at vertex $v$ ($v')$, is defined as
\begin{equation}
    S_e \left( \gamma_{vv'} \right) \equiv \prod_{l \in \gamma_{vv'}} Q_{v_l v_l'}^\dagger
\end{equation}
where the product is over all links $l$ in $\gamma_{vv'}$, $v_l$ and $v_l'$ are the two vertices in $l$, ordered according to the orientation of $\gamma_{vv'}$ (see \figref{fig:example_elec_string}). The magnetic string operator $S_m$ is defined by an open, oriented string on the dual lattice and creates a magnetic vortex-antivortex pair on plaquettes at the endpoints of $\tilde{\gamma}_{pp'}$
\begin{equation}
    S_m \left( \tilde{\gamma}_{vv'} \right) \equiv \prod_{l \perp \tilde{\gamma}_{pp'}} P_{v_l v_l'}^\dagger
\end{equation}
where the product is over all links $l$ that cross $\tilde{\gamma}$ and $v_l$ ($v_l'$) is the vertex to the left (right) of the string as $\tilde{\gamma}$ is traversed from $p$ to $p'$. 
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
    \centering
    \includegraphics[width=0.8\linewidth]{example_elec_string.pdf}
    \caption{ ({\it left}) An example of a charge free $Z_3$ electric string configuration. The electric string operator $S_\gamma^e$ represented by the thick gray line, oriented from the red to blue vertex. ({\it right}) An example of a $Z_3$ electric string configuration with a $+1$ ($-1$) electric charge at the red (blue) vertex; this string configuration results form the action of $S_\gamma^e$ on the left panel.  }
    \label{fig:example_elec_string}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}
%    \centering
%    \includegraphics[width=1.0\columnwidth]{spatial_cors_perp.pdf}
%    \caption{Correlation function between perpendicular dimers on the same (blue) and different (red) 
%    sub-lattice for a $L=64$ lattice.}
%    \label{fig:spatial_dmr_cor}
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%

@@

To test for the possibility of translational and rotational symmetry breaking corresponding to the features in $S_d$ at the Brillouin zone edges, we define several order parameters which would capture such potential symmetry breaking. For rotational symmetry breaking we define 
\begin{equation}
\Mrot \equiv  \frac{2}{L^2} \left(N_h - N_v\right) \label{eq:Mrot},
\end{equation}
where $N_v$ ($N_h$) is the total number of vertical (horizontal) dimers; note that $\exval{M_{\rm{rot}}^2 } \neq 0$ implies there is rotational symmetry breaking. to To capture translational symmetry breaking by one lattice site in a direction perpendicular to the majority of the dimers we define $\Mperp$ 
\begin{align}
\Mperp \equiv  \frac{8}{L^2} \sum_{\Box} \Bigl(  &(-1)^{r_x} \theta\left(\Mrot \right) \Pi_\Box^v \notag\\
+ &(-1)^{r_y}\theta\left(-\Mrot \right) \Pi_\Box^h\Bigr) \label{eq:Mperp},
\end{align}
where $\theta(x)$ is the Heaviside function, and $\Pi_\Box^h$ ($\Pi_\Box^v$) is one if the plaquette $\Box$ has two horizontal (vertical) dimers, and zero otherwise. Similarly, we define $\Mpar$ to be given by \Eqref{eq:Mperp} with $r_x$ and $r_y$ swapped between terms. Additionally, we define order parameters that indicate ``columnar" and ``staggered" dimer order~\cite{Syljuasen2005}:
\begin{align}
\MstagSq \equiv  \frac{1}{L^2} \biggl[ &\biggl( \sum_{\ell_h} (-1)^{r_x+r_y} n_d \left(\vec{r}_{\ell_h}\right) \biggr)^2  \notag\\
+&\biggl( \sum_{\ell_v} (-1)^{r_x+r_y} n_d \left(\vec{r}_{\ell_v}\right) \biggr)^2\biggr]\label{eq:Mstag},
\end{align}
\begin{align}
\McolSq \equiv  \frac{1}{L^2} \biggl[  &\biggl( \sum_{\ell_h} (-1)^{r_x} n_d \left(\vec{r}_{\ell_h}\right) \biggr)^2 \notag \\
+ &\biggl( \sum_{\ell_v} (-1)^{r_y} n_d \left(\vec{r}_{\ell_v}\right) \biggr)^2  \biggr]\label{eq:Mcol},
\end{align}
where the sums over $\ell_h$ ($\ell_v$) are taken over all horizontal (vertical) links. In \figref{fig:order_params} we have plotted the finite-size scaling of these order parameters; we find in all cases that while finite size systems display symmetry breaking due to finite-size effects, these order parameters all vanish in the thermodynamic limit, as demonstrated by the extrapolation in \figref{fig:order_params}.
%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[]
    \centering
    \includegraphics[width=1.0\linewidth]{order_params_plot.pdf}
    \caption{The symmetry breaking order parameters.}
    \label{fig:order_params}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
